### 2012-03-03 userguide/guide.tex
init thesis.tex guide.tex

\documentclass{article}
\usepackage{amsmath,amsthm,amssymb,fullpage,yfonts,graphicx,proof,subfig,wrapfig,appendix,hyperref,mdwlist,wasysym}
\usepackage{upgreek}
\usepackage{epsfig}
\usepackage[bottom]{footmisc}

\begin{document}
\captionsetup{width=.75\textwidth,font=small,labelfont=bf}
\title{\bf Landslide: \\ Systematic Dynamic Race Detection in Kernel-space \\ User Guide}
\author{Benjamin Blum (\textbf{bblum})}
\maketitle

\newcommand\true{\;\textit{true}}
\newcommand\false{\;\textit{false}}

\newcommand\alpher\alpha
\newcommand\beter\beta
\newcommand\gammer\gamma
\newcommand\delter\delta
\newcommand\zeter\zeta
\newcommand\Sigmer\Sigma

\newcommand\NN{\mathbb{N}}
\newcommand\QQ{\mathbb{Q}}
\newcommand\RR{\mathbb{R}}
\newcommand\ZZ{\mathbb{Z}}

\begin{abstract}

\end{abstract}

\section{Introduction}

Race conditions are notoriously difficult to debug. Because of their nondeterministic nature, they often do not manifest at all during testing, and when they do manifest, it is difficult to reproduce

\end{document}