### 2012-05-01 using.tex
buncha minor edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Using Landslide}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:using}
% TODO this whole section

\subsection{Kernel Requirements}
\label{sec:using-requirements}

\subsubsection{Scheduler Functionality}
\label{sec:using-requirements-sched}
% talk about different VM mappings when tracking shm conflicts

\subsection{Instrumenting Kernels with Landslide}

Here we describe the interface by which Landslide understands the execution of the guest kernel.

\subsubsection{In-Kernel Annotations}
\label{sec:using-annotations}

\subsubsection{Instrumenting Within Landslide}
\label{sec:using-student-c}

\subsection{Configuring Landslide's Behaviour}
\label{sec:using-customise}
\subsubsection{Decision Points}
\label{sec:using-decision}
\subsubsection{Search Parameters}
\label{sec:using-search}

\subsection{Test Cases}

We provide a set of test case programs for use with Landslide
% TODO: code listings in appendix

\subsubsection{Landslide-Friendly Test Characteristics}

\subsection{Interpreting Landslide's Results}
\label{sec:using-interpreting}