### 2012-05-01 titlepage.tex

%initialize page style, so contents come out right (see bot) -mjz
\pagestyle{empty}

\title{\bf Landslide: \\
  Systematic Dynamic Race Detection in Kernel-space}
\author{Ben Blum}
\date{May 2012}
\Year{2012}
\trnumber{CMU-CS-12-118}

\committee{
Garth Gibson, Chair \\
Dave Eckhardt
}

\support{This research was sponsored by the U.S. Army Research Office under
grant number W911NF0910273.}

\disclaimer{The views and conclusions contained in
this document are those of the author and should not be interpreted as
representing the official policies, either expressed or implied, of
any sponsoring institution, the U.S. government or any other entity.}

% copyright notice generated automatically from Year and author.
% permission added if \permission{} given.

\keywords{FOO, BAR, Computers, Science}

\maketitle

\begin{dedication}
%There is a wondrous fractal nature to all aspects of the universe.

At the time of writing, my life is focused entirely on this one project. There is a vast amount of thought in this thesis, and many more rich worlds lie hidden in each of the open avenues for future work.

Yet it pales in comparison to the scope of a Ph.D. thesis, and even more so next to all the details of a person's entire life.
After I finish this project (whatever that may mean), it will become but a fond memory, a mere building block in a city of even grander things.

Human nature is beautiful in its ability to exist at any level of perspective, to hold in awareness a story of any size, and to zoom in and out at will to see how multiple parts fit together.
Tearing ourselves up over minutiae is part of the fun, and contemplating overarching directions beyond our ability to change is also part of the fun.

Our true home, though, lies somewhere in between; a place of potential, and of peace. May we always remember to return to it.
\end{dedication}

\pagestyle{plain} % for toc, was empty