### 2012-05-03 conclusion.tex
improve conclusion

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\s}{\fontfamily{jkpvos}\selectfont{}s}

% TODO write this for real
And lo, the Author did{\s}t pre{\s}ent Land{\s}lide to yon re{\s}earch Community, and to the {\s}tudents of XV-CDX as well. Verily, the Kernel developers young and olde found Bugges of Concurren{\s}y in their code. The {\s}oftware then{\s}eforth worked as Intended, cra{\s}hing nevermore, and there was much rejoicing.