### 2012-05-12 thesis.tex

\documentclass[12pt]{book}
%\usepackage{amsmath,amsthm,amssymb,fullpage,yfonts,graphicx,proof,subfig,wrapfig,appendix,hyperref,mdwlist,wasysym}
\usepackage{amsmath,amsthm,amssymb,fullpage,yfonts,graphicx,proof,appendix,hyperref,mdwlist,wasysym}
\usepackage{upgreek}
%\usepackage{times}
\usepackage[charter]{mathdesign}
\usepackage{hyperref}
\usepackage{draftwatermark}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage[usenames,dvipsnames]{xcolor}
%\usepackage{epsfig}
\usepackage[bottom]{footmisc}
\usepackage{mjz-titlepage}
\usepackage{setspace}
\usepackage{subfig}
\setstretch{1.05}

\begin{document}
%\captionsetup{width=.75\textwidth,font=small,labelfont=bf}

\input{titlepage}

\newcommand\true{\;\textit{true}}
\newcommand\false{\;\textit{false}}

\newcommand\alpher\alpha
\newcommand\beter\beta
\newcommand\gammer\gamma
\newcommand\delter\delta
\newcommand\zeter\zeta
\newcommand\Sigmer\Sigma

\newcommand\NN{\mathbb{N}}
\newcommand\QQ{\mathbb{Q}}
\newcommand\RR{\mathbb{R}}
\newcommand\ZZ{\mathbb{Z}}

\begin{abstract}
Systematic exploration is an approach to finding race conditions by deterministically executing every possible interleaving of thread transitions and identifying which ones expose bugs.
Current systematic exploration techniques are suitable for testing user-space programs, but are inadequate for testing kernels, where the testing framework's control over concurrency is more complicated.

We present Landslide, a systematic exploration tool for finding races in kernels.
Landslide targets Pebbles, the kernel specification that students implement in the undergraduate Operating Systems course at Carnegie Mellon University (15-410).
We discuss the techniques Landslide uses to address the general challenges of kernel-level concurrency, and we evaluate its effectiveness and usability as a debugging aid.
We show that our techniques make systematic testing in kernel-space feasible, and that Landslide is a useful tool for doing so in the context of 15-410.
\end{abstract}

\input{acknowledgements}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Race conditions are notoriously difficult to debug.
Because of their nondeterministic nature, they frequently do not manifest at all during testing, and when they do manifest, it is difficult to reproduce them reliably enough to collect enough information to help debugging.

In kernel space, race condition debugging becomes even more difficult. Many aspects of the concurrency implementation itself are part of the system being tested, and may themselves have race-inducing flaws.

% TODO: talk about systematic exploration
%It provides a better alternative to conventional long-running stress tests, because it is less likely to overlook buggy execution patterns, and it enables a testing framework to report more thorough debugging information.

% TODO: make sure to talk about TAs

Landslide\footnote{
Landslide {\em (n)} - A phenomenon which demonstrates that Pebbles are not as stable as you might think.}
is an effort to make easier the process of debugging kernel-space races.
It is geared towards kernels that implement the Pebbles specification, the main project in Operating System Design and Implementation (15-410) at CMU, and implemented as a module for Wind River Simics\textsuperscript{\texttrademark}, the x86 simulator that students use to run their Pebbles kernels.
During execution of a kernel, Landslide records important actions performed by the kernel, attempts to decide at which points in the kernel's execution a preemption will be most likely to expose a bug, and then exercises all possible interleavings of kernel threads around such points.
When Landslide finds a bug (predicted by a set of common checks and heuristics), it stops execution and prints information about the sequence of interleavings that exposed the bug.

With Landslide, we see testing a kernel as a process of manipulating test parameters in two ways: first, in the choice of test case (the user-space program that exercises a specific set of system calls), and second, in the configuration of Landslide in regard to which parts of the kernel are ``interesting'' in the behaviour of the test case and which are irrelevant.
Searching for and understanding race conditions exposed by a given test becomes a joint effort between the programmer and Landslide, combining the programmer's specific knowledge about the design of the kernel and Landslide's ability to explore many interleavings efficiently.

% TODO: this paragraph is incongruous in light of the above three; what audience are you writing for, anyway?
Many techniques exist for dynamic testing of concurrent systems for race conditions.
Systematic exploration, the strategy that Landslide uses, involves modelling a concurrent test case as a series of {\em transitions} by each of a set of {\em agents}
\footnote{In this work, we equate agents with a kernel's scheduler threads.}
between {\em decision points} (i.e., actions which the tool and/or the user determine to be relevant to the concurrent behaviour of the test), and attempting to exercise all possible orderings of the transitions in hope that one or more ordering will expose a bug.

In this work, we present Landslide as a tool and framework for finding race conditions in kernels using systematic exploration, implemented as a module for the Simics simulator and geared towards Pebbles, the specification implemented by students in 15-410 at CMU.
We study the challenges inherent in applying systematic testing techniques in kernel-space, in contrast with user-space applications, and present techniques (some sound, and some heuristic) for addressing them.
% We study the relationship between the developer and the tool during the debugging process, in terms of what information the tool needs the developer to provide beforehand in order to provide meaningful test results.
% Compared to conventional race condition debugging, we show that the relationship Landslide provides makes it a better strategy.
We evaluate Landslide's usefulness in helping students of 15-410 better understand and solve the concurrency problems while implementing a Pebbles kernel, and also as a testing framework for TAs to use while grading student submissions.
We show that Landslide is an effective tool for both purposes.

% TODO: contribution and organisation sections

\input{related} % COMPLETE!
\input{challenges} % COMPLETE!
\input{terminology} % almost complete
\input{design}
\input{using}
\input{evaluation}
\input{futurework}
\input{conclusion}

% TODO: read presubmit.txt and check everything off

\bibliography{citations}{}
\bibliographystyle{alpha}

\appendix
\input{testcases}

\end{document}