### 2012-05-16 acknowledgements.tex
fix the infinite loop thing; add figures

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Save this for last :)

Thanks to Deb Cavlovich, for holding on to this thesis for a few days more for Ben to finish it properly.

Thanks to my advisor, too, I guess, and maybe also to the people I worked with and my friends as well.