### 2012-05-18 acknowledgements.tex
final ack

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter*{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This has been an immensely fun and fulfilling project, and it wouldn't have been possible without the contributions, guidance, and support of many people.

\subsubsection{Big Thanks}

First of all, I owe the lion's share of my gratitude equally to Garth Gibson, Jiri Simsa, and David Eckhardt.

Garth's advisorship over the past year has been excellent, from showing me how to keep my research both useful and personally satisfying, to teaching me how to share ideas with fellow researchers, to helping me think through far-off future work directions. Not working directly with his main research groups was sometimes unsettling, but validation was not in short supply: for example, it was definitely a high point when Garth emailed me unprompted asking for my input on the future of my research field.

Jiri's mentorship was invaluable, both in explanations of complicated algorithms and in general advice on good researcher habits. I would have had a much harder time understanding Dynamic Parial Order Reduction without him, and I suspect my posters and talks would have been much less put-together as well.

Working with David to share my project with the students of 15-410 really made the thesis come together. I knew I would have a rough time making Landslide available without compromising the class's integrity policies, and David helped me make it work. His advice was also crucial when designing the lecture I gave to the class to explain Landslide and recruit students for the user study; I learned more about lecture design during those two weeks than I ever thought there was to know.

\subsubsection{Landslide's Users}

I'm also highly grateful for all the patient work put in by the kernel programmers who used Landslide.

Nathaniel Filardo, as the first person to use Landslide besides myself, ran into several compatibility issues (and outright bugs) in Landslide when instrumenting his kernel. For all Landslide gave him a hard time, though, he stayed with me, was patient while I debugged, and even debugged some on his own when I was too busy myself. It was when he got Landslide working with his kernel that I finally found confidence that it'd be able to work on arbitrary student kernels.

Thanks also to the other members (and former members) of 15-410 staff who used Landslide on their kernels to help me prepare for working with the students: Josiah Boning, Alex Crichton, and Michael Sullivan.

A special form of thanks to the students I worked with, who not only had to navigate Landslide's interfaces but also had to learn the underlying concepts as well, all while under time pressure to finish their kernel projects at the same time. To Nadim Taha, Margaret Schervish, Timothy Passaro, Joon-Sup Han, Pranjal Jumde, and others who did not ask to be mentioned by name: I could not have evaluated Landslide without you.

Another piece of thanks goes to Nadim Taha for his above-and-beyond investigation of the decision trees Landslide explored using his kernel, which resulted in the insights presented in Section~\ref{sec:future-nadim}.

\subsubsection{Collaboration}

Thanks to Wind River for Simics\textsuperscript{\texttrademark}, the simulation software for which Landslide is built. While developing code to fit into larger infrastructure always has its hairy points, Simics's interfaces were generally pleasant, and the expressive power of operating in its simulation environment was invaluable.

Thanks to the members and companies of the PDL Consortium (including Actifo,
APC, EMC, Emulex, Facebook, Fusion-io, Google, Hewlett-Packard, Hitachi, Huawei
Technologies, Intel, Microsoft, NEC, NetApp, Oracle, Panasas, Riverbed,
Samsung, Seagate, STEC, Symantec, VMware, and Western Digital) for their
interest, insights, feedback, and support.
It has been a pleasure to work in this research group.

\subsubsection{Support}

More gratitude goes to the people who helped me revise my thesis document and defence talk. While Garth and David provided the most detailed criticism, many of my friends helped me polish my work: Wolfgang Richter, Carlo Angiuli, Joshua Wise, Ryan Pearl, and Michael Sullivan critiqued this document, and Jiri Simsa, Matthew Maurer, Greg Hanneman, and Carlo Angiuli critiqued my defence presentation.

Finally, huge thanks to Deborah Cavlovich and Catherine Copetas, whose organisational efforts carried me through the Master's program. These are the people who make the gears turn under the hood.

\vspace{0.5in}

\noindent I hope you enjoy reading some or all of this.